### 2021-07-13 article/article.tex
Merged more of James's suggestions

\section{Title}
spinsim: a GPU optimised simulator of spin-half and spin-one quantum systems

@@

\section{Abstract}
	The \emph{spinsim} \emph{python} package simulates spin-half and spin-one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda}, which is an \emph{LLVM} (Low Level Virtual Machine) compiler for \emph{Nvidia Cuda} compatible systems using GPU parallelisation. 
	Along with other optimisations, this allows for speed improvements of up to four orders of magnitude while keeping staying just as accurate.
	\emph{spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use-case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin-half or spin-one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin-one systems of nitrogen vacancy centres (NVCs) and BECs.

@@

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing\cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin-half (ie spin quantum number of $ \frac12 $, also referred to as a qubit), is quantised into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin-one quantum system (ie spin quantum number of $ 1 $, also referred to as a quitrit), which is quantised into three quantum spin levels.

@@

	% \texttt{AtomicPy}\cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package is based, and standard differential equation solvers (such as \emph{Mathematica}'s function \texttt{NDSolve}) to solve the Schroedinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin-half for a faster execution time, at the cost of modelling all effects of the system.
	However, these methods are not completely optimised for our use-case, and therefore come with some issues.

@@

	Given the recent boom in machine learning, many research computers are now equipped with advanced graphics processing units (GPUs).
	Their many cores, which can range from hundreds to tens of thousands in number, allows them to run some highly parallel algorithms much faster than a central processing unit (CPU).
	Examples include calculating colours for many pixels on a screen in the titular graphics processing, or the weights in a large neural network.
	By finding ways to parallelise the problem of solving quantum spin systems, we can use the many cores of a GPU to our advantage in this context as well.

\section{Implementation and architecture}
	% \subsection{Mathematical methods}
	\subsection{Quantum mechanics background}
		The \emph{spinsim} package solves the time-dependent Schr\"{o}dinger equation
		\begin{align}
			i\hbar\frac{\mathrm{d}\psi(t)}{\mathrm{d}t} &= H(t)\psi(t),\label{eq:schroedinger}
		\end{align}
		where the quantum state $ \psi(t) \in \mathbb{C}^N $ is assumed normalised, and the Hamiltonian $ H(t) \in \mathbb{C}^{N \times N} $ is Hermitian.
		Here $ N $ is the number of levels in the quantum system.
		Often we are considering systems with spin of half or one.
		\supervisordisagreement{
			Spin-half is the $ N = 2 $ case, and spin-one is $ N = 3 $.
		}{
		Spin-half is the $ N = 2 $ case, and spin-one is the $ N = 3 $ case.
		}
		We set $ \hbar = 1 $, so that the Hamiltonian has physical dimension of angular frequency.
		% Furthermore, we have chosen for the Hamiltonian $ H(t) $ to have the physical dimension of frequency, rather than the more typical dimension of energy. To elaborate, we represent energies $ E $ of the system as their equivalent frequencies $ f $ from the Planck relation

@@

		% where $ \hbar $ is (reduced) Planck's constant (or equivalently, setting 
		% $ \hbar = 1 $). In addition, rather than being represented by standard coordinates in $ \mathbb{C}^{N \times N} $, in \emph{spinsim} the Hamiltonian $ H(t) $ is instead represented with respect to a choice of basis for the corresponding Lie Algebra, $ \mathfrak{su}(N) $. For example, when set to spin-half mode, the \emph{spinsim} package solves the time dependent Schroedinger equation of the form

		Rather than parametrizing the problem in terms of the matrix elements of $ H(t) $, we consider time varying real coefficients in a linear combination of fixed operators,
		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} \omega_j(t) \mathcal{A}_j.
		\end{align}
		Here we exclude the identity so that the Hamiltonian is traceless, which corresponds to choosing a physically meaningless energy zero point.
		It is well known that a charged spin-half system with magnetic moment $ \overrightarrow{\mu} $, and gyromagnetic ratio $ \gamma $, in a magnetic field $ \overrightarrow{B}(t) $, has Hamiltonian
		\begin{align}
			H(t) &= -\overrightarrow{B}(t)\cdot \overrightarrow{\mu}\\
			% &= -\gamma \overrightarrow{B}(t)\cdot \overrightarrow{J}\\
			&= -\gamma \left(B_x(t) J_x + B_y(t) J_y + B_z(t) J_z\right)\\
			% &= \left(-\gamma B_x(t) J_x\right) + \left(-\gamma B_y(t) J_y\right) + \left(-\gamma B_z(t) J_z\right)\\
			&= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z.
		\end{align}
		\supervisordisagreement{
			Here $ J_x, J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices halved.
		}{
			Here $ J_x, J_y $ and $ J_z $ are spin operators
			\begin{align}
				J_x &= \frac12\begin{pmatrix}
					0 & 1 \\
					1 & 0
				\end{pmatrix},
				&J_y &= \frac12\begin{pmatrix}
					0 & -i \\
					i &  0
				\end{pmatrix},
				&\textrm{and }J_z &= \frac12\begin{pmatrix}
					1 &  0 \\
					0 & -1
				\end{pmatrix},\label{eq:spin_half_operators}
			\end{align}
			equal to the Pauli matrices halved.
		}
		Because of the equivalence of the magnetic field components $ B_j(t) $ and the $ \omega_j(t) $, we henceforth refer to the latter as field functions.

		In the spin-one case there is no standard basis of operators $ \mathcal{A}_j $.
		Choices include the Gell-Mann matrices\cite{gell-mann_symmetries_1962}, and multiple dipole-quadrupole bases\cite{hamley_spin-nematic_2012}\cite{di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra $ \mathfrak{su}(3) $, which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group $ SU(3) $) in the spin-one system.
		We focus on a particular subfamily of of spin-one systems for which the Hamiltonian is a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x, J_y $ and $ J_z $, and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,\todo{Describe this subspace better}
		% With this in mind, we choose to represent the Hamiltonian a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x, J_y $ and $ J_z $, and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q.
		\end{align}

		Note that $ Q $ is proportional to $ Q_{zz} $\cite{hamley_spin-nematic_2012} and $ Q_0 $\cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		The $ J_x $, $ J_y $, $ J_z $ and $ Q $ are the only operators necessary to simulate many spin-one quantum systems, including quadratic Zeeman splitting described by the Breit-Rabi formula\cite{mockler_atomic_1961} important to experiments in our lab.
		The \emph{spinsim} simulator can also be configured to solve a general spin-one system by setting the Hamiltonian to an arbitrary point in $ \mathfrak{su}(3) $ using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) =& \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{u1}(t) U_1 + \omega_{u2}(t) U_2 + \omega_{v1}(t) V_1 + \omega_{v2}(t) V_2.
		\end{align}
		Here the additional operators are those defined in\cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the $ U_1, U_2, V_1 $ and $ V_2 $ operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% Similarly, when \emph{spinsim} is set to spin-one mode, it can solve the Schroedinger equation of the form

@@

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		As well as integrating the Schr\"odinger equation, \emph{spinsim} also has the functionality to calculate the expected spin projection $ \langle \overrightarrow{J}\rangle (t) $ of a system from its state.
		In an experimental setting, one cannot measure the value of the state itself, and must instead measure observables such as spin projection.
		If this is done for an ensemble of systems, the average spin projection over all systems will converge to the expected value as calculated here.
		As an alternative to solving the Schr\"odinger equation, if one is only interested in the dynamics of the expected spin projection, one can instead solve for $ \langle \overrightarrow{J}\rangle (t) $ directly using the Bloch equations\cite{bloch_nuclear_1946}.
		These are three coupled real equations as compared to three coupled complex equations for the spin-one Schr\"odinger equation.
		However, the Bloch equations cannot be used to compute $ \psi(t) $, and can only fully model spin-half systems (and spin-one systems approximated to spin-half).
		Thus, with \emph{spinsim} we choose to solve the full Schr\"odinger equation instead.

@@

		% where $ \cdot^\dagger $ is the adjoint operator. The expected spin projection can be interpreted as the average direction that the spin system is oriented in space when many systems of the equivalent state are measured (due to the Heisenberg uncertainty principle, only one component of the exact orientation of a quantum state can be known at any point in time, which is why we need to deal with expected values rather than absolute values). However, it can also be used to model the overall orientation of an ensemble of many quantum systems, like a BEC, for instance. \emph{spinsim} has the functionality to calculate the expected spin projection of a system from its state.

	\subsection{Unitary time evolution and the Magnus expansion}
		Since $H(t)$ is Hermitian, it follows from Equation \eqref{eq:schroedinger} that $\|\psi(t)\|^2 = \|\psi(t_0)\|^2$ for all $t$.
		Therefore it is possible to write $ \psi(t) $ in terms of a unitary transformation $ \mathcal{U}(t, t_0) $ of the state $ \psi(t_0) $, for any time $ t_0 $, that is, $ \psi(t) = \mathcal{U}(t, t_0)\psi(t_0) $.
		It then follows from Equation \eqref{eq:schroedinger} that $ \mathcal{U}(t, t_0) $ also follows a Schr\"{o}dinger equation,
		\begin{align}
			i\hbar\frac{\mathrm{d}\mathcal{U}(t, t_0)}{\mathrm{d}t} &= H(t)\mathcal{U}(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation \eqref{eq:schroedinger} for $ \psi(t) $ given a particular $ \psi(t_0) $, one only needs to solve Equation \eqref{eq:unitary_schroedinger} for $ \mathcal{U}(t, t_0) $.
		% Note that because $ \mathcal{U}(t, t_0) $ is unitary, the calculated value of $ \psi(t) $ using this method will not change in magnitude, conserving probability.
		Since $ \mathcal{U}(t, t_0) $ is unitary, it can be written as a matrix exponential of a anti-Hermitian (also termed skew-Hermitian) matrix.
		If the Hamiltonian is a of constant value $ H(t) = \mathcal{H} $, then this exponential is \note{Should we use H, ie no argument, or H(t'), ie arbitrary argument, to represent constancy?} % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i (t - t_0) \mathcal{H}).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series\cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for $ \mathcal{U}(t, t_0) $ is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The well known Dyson series\cite{kalev_integral-free_2020} gives an explicit expression for $ \mathcal{U}(t, t_0) $ in terms of multiple time integrals over nested time commutators of $ H(t') $.
		While the Dyson series has recently been used numerically\cite{kalev_integral-free_2020}, it is challenging to work with because once truncated, it is in general no longer unitary.

		The Magnus series, in contrast, gives $ \mathcal{U}(t, t_0) $ in terms of an exponential of a series of anti-Hermitian operators, specifically
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp\left(\Omega(t, t_0)\right) = \exp\left(\sum_{m = 1}^\infty \Omega_m(t, t_0)\right).
		\end{align}
		As such, the Magnus series explicitly preserves unitarity when truncated\cite{magnus_exponential_1954}.
		The terms in the Magnus series are integrals over nested time commutators of the Hermitian matrix $ A(t') $, where, in the case of quantum mechanics, $ A(t') = -iH(t') $.
		For instance, the first three terms are\cite{blanes_magnus_2009}
		\begin{align}
			\Omega_1(t, t_0) &= \int_0^t\mathrm{d}t_1A(t_1)\\
			\Omega_2(t, t_0) &= \frac12\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2[A(t_1), A(t_2)]\\
			\Omega_3(t, t_0) &= \frac16\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2\int_{t_0}^{t_2}\mathrm{d}t_3\left([A(t_1), [A(t_2), A(t_3)]] + [A(t_3), [A(t_2), A(t_1)]]\right),
		\end{align}
		where $ [X, Y] = XY - YX $ is the commutator.
		Note that, when truncated to first order, the Magnus expansion $ \mathcal{U}(t, t_0) = \exp\left(\int_{t_0}^t\mathrm{d}t_1\left(-iH(t_1)\right)\right) $ reduces to Equation \eqref{eq:exp_sol_of_constant}, with $ H(t) $ approximated by its average value over the interval $ [t_0, t] $.

@@

	\subsection{Lie-Trotter based Exponentiator}
		Evaluating matrix exponentials is a core part of the integration algorithm.
		Rather than exponentiating the Hamiltonian directly as in Equation \eqref{eq:cf4_implementation}, \emph{spinsim} works with the field functions $ \omega_j(t) $.
		% as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array.
		% This allows for matrix exponentiators that take advantage of this representation.
		% For all implementations of exponentiation, the \emph{spinsim} computes the matrix exponential
		% % \begin{align}
		% % 	\exp(-iHt) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
		% % 	&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		% % \end{align}
		% \begin{align}
		% 	\exp(-iHt) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
		% 	&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		% \end{align}
		% In practice, $ \omega $ is the Lie algebra (spin and quadrature) basis representation of $ \overline{H_1} $ or $ \overline{H_2} $ defined in Equations \eqref{eq:cf4_sample_1} and \eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation \eqref{eq:cf4_implementation}.

>>>>

\section{Title}
Spinsim: a GPU optimized simulator of spin-half and spin-one quantum systems

@@

\section{Abstract}
	The \emph{Spinsim} \emph{python} package simulates spin-half and spin-one quantum mechanical systems following a time dependent Shroedinger equation.
	It makes use of \texttt{numba.cuda}, which is an \emph{LLVM} (Low Level Virtual Machine) compiler for \emph{Nvidia Cuda} compatible systems using GPU parallelisation. 
	Along with other optimisations, this allows for speed improvements of up to four orders of magnitude while keeping staying just as accurate.
	\emph{Spinsim} is available for installation on \emph{PyPI}, and the source code is available on \emph{github}.
	The initial use-case for the package will be to simulate quantum sensing-based Bose Einstein condensate (BEC) experiments for the Monash University School of Physics and Astronomy spinor BEC lab, but we anticipate it will be useful in simulating any range of spin-half or spin-one quantum systems with time dependent Hamiltonians that cannot be solved analytically.
	These appear in the fields of nuclear magnetic resonance (NMR), nuclear quadrupole resonance (NQR) and magnetic resonance imaging (MRI) experiments and quantum sensing, and with the spin-one systems of nitrogen vacancy centres (NVCs) and BECs.

@@

\section{Introduction}
\subsection{Motivation}
	Ultracold rubidium atoms have proven their effectiveness in state of the art technologies in quantum sensing\cite{degen_quantum_2017}, the use of quantum mechanics to make precise measurements of small signals.
	The rotation of these atoms can be modelled as quantum spin systems, which is the quantum mechanical model for objects with angular momentum.
	The simplest spin system, spin-half (ie spin quantum number of $ \frac12 $, also referred to as a qubit), is quantized into just two quantum spin levels, and this describes the motion of some fundamental particles such as electrons.
	However, systems more practical for sensing, such as ultracold rubidium atoms, are more accurately described as a spin-one quantum system (ie spin quantum number of $ 1 $, also referred to as a quitrit), which is quantized into three quantum spin levels.

@@

	% \texttt{AtomicPy}\cite{morris_qcmonkatomicpy_2018}
	In the past, the spinor Bose Einstein condensate (spinor BEC) lab at Monash University used an in-house, \emph{cython} based script, on which this package is based, and standard differential equation solvers (such as \emph{Mathematica}'s function \texttt{NDSolve}) to solve the Schroedinger equation for quantum sensing spin systems.
	Spin one systems are sometimes approximated to spin-half for a faster execution time, at the cost of modelling all effects of the system.
	However, these methods are not completely optimized for our use-case, and therefore come with some issues.

@@

	Given the recent boom in machine learning, many research computers are now equipped with advanced graphics processing units (GPUs).
	Their many cores, which can range from hundreds to tens of thousands in number, allows them to run some highly parallel algorithms much faster than a central processing unit (CPU).
	Examples include calculating colours for many pixels on a screen in the titular graphics processing, or the weights in a large neural network.
	By finding ways to parallelize the problem of solving quantum spin systems, we can use the many cores of a GPU to our advantage in this context as well.

\section{Implementation and architecture}
	% \subsection{Mathematical methods}
	\subsection{Quantum mechanics background}
		The \emph{Spinsim} package solves the time-dependent Schr\"{o}dinger equation
		\begin{align}
			i\hbar\frac{\mathrm{d}\psi(t)}{\mathrm{d}t} &= H(t)\psi(t),\label{eq:schroedinger}
		\end{align}
		where the quantum state $ \psi(t) \in \mathbb{C}^N $ is assumed normalized, and the Hamiltonian $ H(t) \in \mathbb{C}^{N \times N} $ is Hermitian.
		Here $ N $ is the number of levels in the quantum system.
		Often we are considering systems with spin of half or one.
		Spin-half is the $ N = 2 $ case, and spin-one is $ N = 3 $.
		% \supervisordisagreement{
		% 	Spin-half is the $ N = 2 $ case, and spin-one is $ N = 3 $.
		% }{
		% Spin-half is the $ N = 2 $ case, and spin-one is the $ N = 3 $ case.
		% }
		We set $ \hbar = 1 $, so that the Hamiltonian has physical dimension of angular frequency.
		% Furthermore, we have chosen for the Hamiltonian $ H(t) $ to have the physical dimension of frequency, rather than the more typical dimension of energy. To elaborate, we represent energies $ E $ of the system as their equivalent frequencies $ f $ from the Planck relation

@@

		% where $ \hbar $ is (reduced) Planck's constant (or equivalently, setting 
		% $ \hbar = 1 $). In addition, rather than being represented by standard coordinates in $ \mathbb{C}^{N \times N} $, in \emph{Spinsim} the Hamiltonian $ H(t) $ is instead represented with respect to a choice of basis for the corresponding Lie Algebra, $ \mathfrak{su}(N) $. For example, when set to spin-half mode, the \emph{Spinsim} package solves the time dependent Schroedinger equation of the form

		Rather than parametrizing the problem in terms of the matrix elements of $ H(t) $, we consider time varying real coefficients in a linear combination of fixed operators,
		\begin{align}
			H(t) &= \sum_{j = 1}^{N^2 - 1} \omega_j(t) \mathcal{A}_j.
		\end{align}
		Here we exclude the identity so that the Hamiltonian is traceless, which corresponds to choosing a physically meaningless energy zero point.
		It is well known that a charged spin-half system with magnetic moment $ \overrightarrow{\mu} $, and gyromagnetic ratio $ \gamma $, in a magnetic field $ \overrightarrow{B}(t) $, has Hamiltonian
		\begin{align}
			H(t) &= -\overrightarrow{B}(t)\cdot \overrightarrow{\mu}\\
			% &= -\gamma \overrightarrow{B}(t)\cdot \overrightarrow{J}\\
			&= -\gamma \left(B_x(t) J_x + B_y(t) J_y + B_z(t) J_z\right)\\
			% &= \left(-\gamma B_x(t) J_x\right) + \left(-\gamma B_y(t) J_y\right) + \left(-\gamma B_z(t) J_z\right)\\
			&= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z.
		\end{align}
		Here $ J_x $, $ J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices\cite[(p169)]{j_j_sakurai_jun_john_modern_2011} halved.
		% \supervisordisagreement{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators, equal to the Pauli matrices halved.
		% }{
		% 	Here $ J_x, J_y $ and $ J_z $ are spin operators
		% 	\begin{align}
		% 		J_x &= \frac12\begin{pmatrix}
		% 			0 & 1 \\
		% 			1 & 0
		% 		\end{pmatrix},
		% 		&J_y &= \frac12\begin{pmatrix}
		% 			0 & -i \\
		% 			i &  0
		% 		\end{pmatrix},
		% 		&\textrm{and }J_z &= \frac12\begin{pmatrix}
		% 			1 &  0 \\
		% 			0 & -1
		% 		\end{pmatrix},\label{eq:spin_half_operators}
		% 	\end{align}
		% 	equal to the Pauli matrices halved.
		% }
		Because of the equivalence of the magnetic field components $ B_j(t) $ and the $ \omega_j(t) $, we henceforth refer to the latter as field functions.

		In the spin-one case there is no standard basis of operators $ \mathcal{A}_j $.
		Choices include the Gell-Mann matrices\cite{gell-mann_symmetries_1962}, and multiple dipole-quadrupole bases\cite{hamley_spin-nematic_2012, di_dipolequadrupole_2010}.
		In general, we can choose any basis from the 8-dimensional Lie algebra $ \mathfrak{su}(3) $, which is the vector space of traceless Hermitian operators that can generate transformations (from the corresponding Lie group $ SU(3) $) in the spin-one system.
		We focus on a particular subfamily of of spin-one systems for which the Hamiltonian is a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x $, $ J_y $ and $ J_z $, (labelled $ L_x $, $ L_y $ and $ L_z $ in Reference~\cite{hamley_spin-nematic_2012}) and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,% \todo{Describe this subspace better}

		% This spin-one exponentiator evolves Hamiltonians spanned by $ J_x $, $ J_y $, $ J_z $ and $ Q $ which is sufficient for three level systems in arbitrary bias fields, but with single-photon coupling.
		% An exponentiator capable of evolving an arbitrary spin-one Hamiltonian, for example with different coupling between the lower and upper pairs of states, or with two-photon coupling, is included in the package.

		% With this in mind, we choose to represent the Hamiltonian a linear combination of matrices from a 4-dimensional subspace of $ \mathfrak{su}(3) $, consisting of the spin matrices $ J_x, J_y $ and $ J_z $, and a single quadrupole operator $ Q = \mathrm{diag}(1, -2, 1)/3 $,
		\begin{align}
			H(t) &= \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q.
		\end{align}

		Note that $ Q $ is proportional to $ Q_{zz} $\cite{hamley_spin-nematic_2012} and $ Q_0 $\cite{di_dipolequadrupole_2010} from alternative quadrupole bases.
		The $ J_x $, $ J_y $, $ J_z $ and $ Q $ are the only operators necessary to simulate many spin-one quantum systems in arbitrary bias fields, but with single-photon coupling.
		This includes quadratic Zeeman splitting described by the Breit-Rabi formula\cite{mockler_atomic_1961} important to experiments in our lab.
		The \emph{Spinsim} simulator can also be configured to solve a general spin-one system by setting the Hamiltonian to an arbitrary point in $ \mathfrak{su}(3) $ using the full quadrupole basis, which extends the possible Hamiltonian to
		\begin{align}
			H(t) =& \omega_x(t) J_x + \omega_y(t) J_y + \omega_z(t) J_z + \omega_q(t) Q\nonumber\\
			&+ \omega_{u1}(t) U_1 + \omega_{u2}(t) U_2 + \omega_{v1}(t) V_1 + \omega_{v2}(t) V_2.
		\end{align}
		Here the additional operators are those defined in\cite{di_dipolequadrupole_2010}. Note that this is included for completeness, but has not been thoroughly tested, as our lab has no physical context for modelling the $ U_1, U_2, V_1 $ and $ V_2 $ operators.
		% In general ... Lie algebra ... mathfrak ... admits a basis ... .
		% \begin{align}
		% 	\frac{\mathrm{d}}{\mathrm{d}t}\psi(t) = -i 2\pi (f_x(t) J_x + f_y(t) J_y + f_z(t) J_z) \psi(t),
		% \end{align}

@@

		% Similarly, when \emph{Spinsim} is set to spin-one mode, it can solve the Schroedinger equation of the form

@@

		% Too far out of scope
		% Alternative: integrate Bloch equations 3 coupled real equations
		% fast and parallel function from wavefunctions
		As well as integrating the Schr\"odinger equation, \emph{Spinsim} also has the functionality to calculate the expected spin projection $ \langle \overrightarrow{J}\rangle (t) $ of a system from its state.
		In an experimental setting, one cannot measure the value of the state itself, and must instead measure observables such as spin projection.
		If this is done for an ensemble of systems, the average spin projection over all systems will converge to the expected value as calculated here.
		As an alternative to solving the Schr\"odinger equation, if one is only interested in the dynamics of the expected spin projection, one can instead solve for $ \langle \overrightarrow{J}\rangle (t) $ directly using the Bloch equations\cite{bloch_nuclear_1946}.
		These are three coupled real equations as compared to three coupled complex equations for the spin-one Schr\"odinger equation.
		However, the Bloch equations cannot be used to compute $ \psi(t) $, and can only fully model spin-half systems (and spin-one systems approximated to spin-half).
		Thus, with \emph{Spinsim} we choose to solve the full Schr\"odinger equation instead.

@@

		% where $ \cdot^\dagger $ is the adjoint operator. The expected spin projection can be interpreted as the average direction that the spin system is oriented in space when many systems of the equivalent state are measured (due to the Heisenberg uncertainty principle, only one component of the exact orientation of a quantum state can be known at any point in time, which is why we need to deal with expected values rather than absolute values). However, it can also be used to model the overall orientation of an ensemble of many quantum systems, like a BEC, for instance. \emph{Spinsim} has the functionality to calculate the expected spin projection of a system from its state.

	\subsection{Unitary time evolution and the Magnus expansion}
		Since $H(t)$ is Hermitian, it follows from Equation~\eqref{eq:schroedinger} that $\|\psi(t)\|^2 = \|\psi(t_0)\|^2$ for all $t$.
		Therefore it is possible to write $ \psi(t) $ in terms of a unitary transformation $ \mathcal{U}(t, t_0) $ of the state $ \psi(t_0) $, for any time $ t_0 $, that is, $ \psi(t) = \mathcal{U}(t, t_0)\psi(t_0) $.
		It then follows from Equation~\eqref{eq:schroedinger} that $ \mathcal{U}(t, t_0) $ also follows a Schr\"{o}dinger equation,
		\begin{align}
			i\hbar\frac{\mathrm{d}\mathcal{U}(t, t_0)}{\mathrm{d}t} &= H(t)\mathcal{U}(t, t_0).\label{eq:unitary_schroedinger}
		\end{align}

		Thus, to solve Equation~\eqref{eq:schroedinger} for $ \psi(t) $ given a particular $ \psi(t_0) $, one only needs to solve Equation~\eqref{eq:unitary_schroedinger} for $ \mathcal{U}(t, t_0) $.
		% Note that because $ \mathcal{U}(t, t_0) $ is unitary, the calculated value of $ \psi(t) $ using this method will not change in magnitude, conserving probability.
		Since $ \mathcal{U}(t, t_0) $ is unitary, it can be written as a matrix exponential of a anti-Hermitian (also termed skew-Hermitian) matrix.
		If the Hamiltonian is a of constant value $ H(t) = \mathcal{H} $, then this exponential is % \note{Should we use H, ie no argument, or H(t'), ie arbitrary argument, to represent constancy?} % Check if continuity requirement on unitary
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp(-i (t - t_0) \mathcal{H}).\label{eq:exp_sol_of_constant}
		\end{align}
		% The act of integrating a linear differential equation by exponentiation can be generalised for time dependent Hamiltonians using the Magnus Series\cite{blanes_magnus_2009}, where
		For a time varying Hamiltonian, the general solution for $ \mathcal{U}(t, t_0) $ is much more complex, because it encapsulates the full solution to the time-dependent Schr\"{o}dinger equation.
		The well known Dyson series\cite{kalev_integral-free_2020} gives an explicit expression for $ \mathcal{U}(t, t_0) $ in terms of multiple time integrals over nested time commutators of $ H(t') $.
		While the Dyson series has recently been used numerically\cite{kalev_integral-free_2020}, it is challenging to work with because once truncated, it is in general no longer unitary.

		The Magnus series, in contrast, gives $ \mathcal{U}(t, t_0) $ in terms of an exponential of a series of anti-Hermitian operators, specifically
		\begin{align}
			\mathcal{U}(t, t_0) &= \exp\left(\Omega(t, t_0)\right) = \exp\left(\sum_{m = 1}^\infty \Omega_m(t, t_0)\right).
		\end{align}
		As such, the Magnus series explicitly preserves unitarity when truncated\cite{magnus_exponential_1954}.
		The terms in the Magnus series are integrals over nested time commutators of the Hermitian matrix $ A(t') $, where, in the case of quantum mechanics, $ A(t') = -iH(t') $.
		For instance, the first three terms are\cite{blanes_magnus_2009}
		\begin{align}
			\Omega_1(t, t_0) &= \int_0^t\mathrm{d}t_1A(t_1)\\
			\Omega_2(t, t_0) &= \frac12\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2[A(t_1), A(t_2)]\\
			\Omega_3(t, t_0) &= \frac16\int_{t_0}^t\mathrm{d}t_1\int_{t_0}^{t_1}\mathrm{d}t_2\int_{t_0}^{t_2}\mathrm{d}t_3\left([A(t_1), [A(t_2), A(t_3)]] + [A(t_3), [A(t_2), A(t_1)]]\right),
		\end{align}
		where $ [X, Y] = XY - YX $ is the commutator.
		Note that, when truncated to first order, the Magnus expansion $ \mathcal{U}(t, t_0) = \exp\left(\int_{t_0}^t\mathrm{d}t_1\left(-iH(t_1)\right)\right) $ reduces to Equation~\eqref{eq:exp_sol_of_constant}, with $ H(t) $ approximated by its average value over the interval $ [t_0, t] $.

@@

	\subsection{Lie-Trotter based Exponentiator}
		Evaluating matrix exponentials is a core part of the integration algorithm.
		Rather than exponentiating the Hamiltonian directly as in Equation~\eqref{eq:cf4_implementation}, \emph{Spinsim} works with the field functions $ \omega_j(t) $.
		% as a vector representation of the Hamiltonian of the spin system being simulated, rather than generic matrix elements in an array.
		% This allows for matrix exponentiators that take advantage of this representation.
		% For all implementations of exponentiation, the \emph{Spinsim} computes the matrix exponential
		% % \begin{align}
		% % 	\exp(-iHt) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
		% % 	&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		% % \end{align}
		% \begin{align}
		% 	\exp(-iHt) &= E(\omega_x, \omega_y, \omega_z, \omega_q)\\
		% 	&= \exp(-i (\omega_x J_x + \omega_y J_y + \omega_z J_z + \omega_q Q)).
		% \end{align}
		% In practice, $ \omega $ is the Lie algebra (spin and quadrature) basis representation of $ \overline{H_1} $ or $ \overline{H_2} $ defined in Equations~\eqref{eq:cf4_sample_1} and~\eqref{eq:cf4_sample_2}, which need to be exponentiated in Equation~\eqref{eq:cf4_implementation}.
